---
title: "Getting Started with atspR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with atspR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
library(atspR)
```

## Overview

`atspR` (**Automated Time-Series Preprocessing in R**) provides a complete,
modular preprocessing framework for time-series data — from raw input to
cross-validated, model-ready datasets.

---

## Installation

```r
devtools::install_github("example/Automated-Time-Series-Preprocessing-in-R")
```

---

## Case 1: Date only (daily data)

```{r case1, eval=FALSE}
library(atspR)

df <- data.frame(
  date  = as.Date(c("2024-01-01","2024-01-02","2024-01-04","2024-01-05")),
  sales = c(100, 120, NA, 130),
  temp  = c(28.1, 27.5, 29.0, 28.8)
)

# 1. Standardise missing values
df <- standardize_na(df, na_strings = c("-", "Missing"))

# 2. Fill missing timestamps
gap <- fill_time_gaps(df, time_col = "date", n = 1, unit = "day")

# 3. Run full pipeline
result <- ts_preprocess(
  data         = gap$data,
  train_ratio  = 0.8,
  scale_method = "minmax",
  target_col   = "sales"
)

result$cv_summary
```

---

## Case 2: Datetime in one column

```{r case2, eval=FALSE}
df$datetime <- as.POSIXct(df$datetime,
                           format = "%Y-%m-%d %H:%M:%S",
                           tz     = "UTC")

gap    <- fill_time_gaps(df, time_col = "datetime", n = 1, unit = "hour")
result <- ts_preprocess(data = gap$data, target_col = "temp")
```

---

## Case 3: Date + Time in separate columns

```{r case3, eval=FALSE}
# combine_datetime รองรับ time_type 3 แบบ:
#   "hour"   → integer 0-23  เช่น 8, 14
#   "hhmm"   → integer HHMM  เช่น 830, 1430
#   "string" → character     เช่น "08:30", "14:30"

df2 <- combine_datetime(raw_data,
                        date_col  = "Date",
                        time_col  = "Time",
                        new_col   = "datetime",
                        time_type = "string",
                        tz        = "Asia/Bangkok")

gap    <- fill_time_gaps(df2, time_col = "datetime", n = 1, unit = "hour")
result <- ts_preprocess(
  data          = gap$data,
  impute_method = "knn",
  scale_method  = "minmax",
  target_col    = "VPD",
  k_folds       = 5L
)
```

---

## Walk-Forward Cross-Validation

`atspR` uses **walk-forward validation** — the correct approach for
time-series. Each fold's validation set always follows its training set
in time.

```
Seed(20%)  Fold1  Fold2  Fold3  Fold4  Fold5
[─────────][─────][─────][─────][─────][─────]

Fold 1: train = seed              → val = fold1
Fold 2: train = seed + fold1      → val = fold2
Fold 3: train = seed + fold1+2    → val = fold3
```

```{r cv_example, eval=FALSE}
data(airquality)
clean <- airquality[complete.cases(airquality), ]
sp    <- split_data(clean, verbose = FALSE)
sc    <- scale_data(sp, method = "minmax", verbose = FALSE)

cv <- cross_validate(
  sc,
  target_col     = "Ozone",
  k              = 5L,
  min_train_size = 0.2
)

cv$summary
cv$fold_results
```

---

## Export Results

```{r export, eval=FALSE}
export_pipeline(result, dir = "output", prefix = "atspR")
# Saves:
#   output/atspR_train_scaled.csv
#   output/atspR_test_scaled.csv
#   output/atspR_data_clean.csv
#   output/atspR_missing_report.csv
#   output/atspR_scale_params.csv
#   output/atspR_cv_summary.csv
```

---

## Available Functions

| Function | Description |
|---|---|
| `standardize_na()` | Convert custom missing indicators to `NA` |
| `coerce_numeric()` | Auto-convert character columns to numeric |
| `combine_datetime()` | Merge date + time columns → `POSIXct` |
| `fill_time_gaps()` | Insert missing timestamps |
| `missing_analysis()` | Analyse & visualise missing values |
| `handle_missing()` | Drop rows or impute (linear / KNN) |
| `visualize_data()` | Scatter plots per variable |
| `split_data()` | Temporal train / test split |
| `scale_data()` | Feature scaling |
| `inverse_scale()` | Reverse scaling to original units |
| `cross_validate()` | Walk-forward k-fold CV |
| `ts_preprocess()` | Run all steps in one call |
| `export_pipeline()` | Export results to CSV |
